---
title: "Import Vpro R format data"
author: "William H MacKenzie"
date: "18/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
library(ggplot2)
require(mapview)
require(sf)
require(stringr)
require(DBI) #loads odbc as well
require(data.table)
source("./_functions/update_taxa.R")
```
# Background

This script imports a VPRO _Veg table in csv format and 
1. compares to the latest provincial standard updates species codes to current provincial standards.
2. flags codes that do not match the provincial code list or are considered excluded
3. identifies any duplicated records and merges into a single record.
4. checks that species are entered in the correct layer by lifeform
5.  updates strata totals where those values are not included
6. adds a layer value to A2 or B2 where only a strata total.

Summary statistic of the collection are then output
The final table should be re-imported into the Vpro BECMaster database



```{r import data}
 becmaster <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=E:/Arctic/Vpro64_Data_Arctic/AVACanada2025.accdb;")


plot.env <- dbReadTable(becmaster, "AVA_CANADA_2025_Env") %>% mutate(Longitude = ifelse(Longitude<0, Longitude, 0-Longitude))
plot.info <- dbReadTable(becmaster, "AVA_CANADA_2025_Admin")
plot.veg <- dbReadTable(becmaster, "AVA_CANADA_2025_Veg") 
dbDisconnect(becmaster)

sppmaster <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=E:/Arctic/ArcticSpeciesInfo/NationalSpeciesLibrary.accdb;")
taxon.all  <- dbReadTable(sppmaster, "USysAllSpecs")
taxon.all1 <- dbReadTable(sppmaster, "AllCodes")
pasl  <- dbReadTable(sppmaster, "PASL_Species2019")
vascan <- dbReadTable(sppmaster, "Vascan_Taxa_v37_11")
arctic  <- dbReadTable(sppmaster, "ArcticSpecies")

dbDisconnect(sppmaster)
missing.cover <- plot.veg %>% filter(is.na(Cover1) & is.na(Cover2) & is.na(Cover3) & is.na(Cover4) & is.na(Cover5) & is.na(Cover6) & is.na(Cover7) & is.na(TotalA) & is.na(TotalB) & is.na(Cover5a)& is.na(Cover5b)& is.na(Cover5c))
missed.spp <- missing.cover %>% select(Species) %>% distinct

missing.totB <- plot.veg %>% filter(is.na(TotalB) & c(!is.na(Cover4) | !is.na(Cover5) | !is.na(Cover5a)| !is.na(Cover5b)| !is.na(Cover5c)))
missing.totB <- missing.totB %>% rowwise %>% mutate(TotalB = sum(Cover4, Cover5, Cover5a, Cover5b, Cover5c, na.rm = T))
 becmaster <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=E:/Arctic/Vpro64_Data_Arctic/AVACanada2025.accdb;")
 dbWriteTable(becmaster, "update_totalb", missing.totB, overwrite = T, batch_rows = 1)
 dbDisconnect(becmaster)
# sppmaster2 <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=E:/Arctic/ArcticSpeciesInfo/NationalSpeciesLibrary.accdb;")
# taxon.all2  <- dbReadTable(sppmaster2, "PASL_Species2019")
# dbDisconnect(sppmaster2)

sppmaster3 <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=E:/Arctic/ArcticSpeciesInfo/NationalSpeciesLibrary.accdb;")
vascan  <- dbReadTable(sppmaster3, "Vascan_Taxa_v37_11")
dbDisconnect(sppmaster3)
#taxon.all <- fread("D:/CommonTables/SpeciesMaster/SpeciesMaster01Dec2020.csv", header = T, stringsAsFactors = F, strip.white = T)
#taxon.info2 <- taxon.all1  %>% mutate(ScientificName = FullTaxaName)#%>% filter(SYNONYM == "FALSE")
taxon.info <- taxon.all %>% filter(Codetype == "U" | Codetype == "D") %>% dplyr::select(-ID)%>% mutate(Species = Code)
vascan.info <- vascan %>% mutate(ScientificName = ifelse(taxonRank == "genus", paste0(genus, " spp."), 
                ifelse(taxonRank == "species", paste0(genus, " ", specificEpithet),
                ifelse(taxonRank == "subspecies", paste0(genus, " ", specificEpithet, " ssp. ", infraspecificEpithet),
                ifelse(taxonRank == "variety", paste0(genus, " ", specificEpithet, " var. ", infraspecificEpithet),
                ifelse(taxonRank == "family", family, NA))))))

## only currently accepted codes
taxon.excluded <- taxon.all %>% filter(Codetype == c("X")) %>% dplyr::select(-ID) %>% mutate(Species = Code)
#taxon_geog_update <- fread("./raw_data/Excluded_Spp_Global_UpdateCodes.csv")  %>% filter(!GlobalUpdate == "")
#taxon_badcode_update <- fread("./raw_data/Bad_Codes_Update.csv") %>% filter(!Code == "", !is.na(Code))
#taxon_badcode_update2 <- fread("./raw_data/Bad_Codes_Update2.csv") %>% filter(!Code == "")
#taxon_badcode_update2 <- fread("./raw_data/Bad_Codes_Update3.csv") %>% filter(!Code == "")
#taxon_badcode_update <- rbind(taxon_badcode_update,taxon_badcode_update2)
```

```{r fix Bliss}
bliss.proj = c("Bliss1984b")
plot.proj <- plot.env %>% select(PlotNumber, FieldNumber, ProjectID, StrataCoverHerb, StrataCoverMoss)
taxon.names <- taxon.info %>% select(Species,ScientificName, Lifeform )
bliss <- left_join(plot.proj, plot.veg) %>% filter(ProjectID %in% bliss.proj) %>% left_join(taxon.names) %>% select(PlotNumber, FieldNumber, StrataCoverHerb, StrataCoverMoss, Species, ScientificName,Lifeform, Cover6, Cover7) %>%  rowwise() %>% mutate(cover = sum(Cover6,Cover7, na.rm=T))
bliss.wide <- bliss %>% pivot_wider(id_cols = c("Species", "ScientificName", "Lifeform"), names_from = "FieldNumber", values_from = cover) %>%  mutate(across(where(is.numeric), round, 3))
#%>% group_by(PlotNumber) %>% mutate(sumcover6 = sum(Cover6, na.rm = TRUE), sumcover7 = sum(Cover7, na.rm = TRUE)) %>% mutate(cov6x = StrataCoverHerb/sumcover6, cov7x = StrataCoverMoss/sumcover7)
#bliss.new <- bliss %>% mutate(Cover6 = Cover6*cov6x, Cover7 = Cover7*cov7x) %>% mutate(across(where(is.numeric), round, 2))
fwrite(bliss.wide, "blissPlantProduction.csv")

```
## import Quebec data
```{r}
quebecmaster <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=E:/Arctic/Vpro64_Data_Arctic/Quebec2025.accdb;")
plot.id <- dbReadTable(quebecmaster, "Quebec_2025_ArcticPlotsID")
quebec.raw <- dbReadTable(quebecmaster,"Quebec_2025_ArcticPlots")
dbDisconnect(quebecmaster)

quebec.env <- quebec.raw %>% select(PlotNumber, NUM_PARC, Longitude, Latitude, Elev, campagne, GR_HABITAT, TYPE_HABITAT)#, QUALIF_PH, QUALIF_DR, QUALIF_MI, SNAME_PRELIM)  %>% distinct 
quebec.env <- left_join(quebec.env, plot.id)%>% mutate(Longitude = ifelse(Longitude<0, Longitude, 0-Longitude))

quebec.veg <- quebec.raw %>% select(PlotNumber, NOM_ESP, COUVERT) %>% mutate(NOM_ESP = str_replace(NOM_ESP, " sp.", " spp."))
quebec.veg <- quebec.veg %>% mutate(NOM_ESP= str_replace(NOM_ESP,  "subsp.", "ssp."))
setDT(quebec.veg)[setDT(taxon.all), `:=` ("spp" = Code, "layer" = Lifeform), on = c("NOM_ESP" = "ScientificName")]
# missing <- quebec.veg %>% filter(is.na(spp)) %>% select(NOM_ESP, spp) %>% distinct
# fwrite(missing, "./out_data/missing_QC_spp.csv")
spp.fixed <- fread("./out_data/missing_QC_spp_fixed.csv") %>% filter(!newspp == "")
setDT(quebec.veg)[setDT(spp.fixed), "spp" := newspp, on = c("NOM_ESP" = "NOM_ESP")]
setDT(quebec.veg)[setDT(taxon.info), `:=` ("layer" = Lifeform), on = c("spp" = "Code")]
 # missing <- quebec.veg %>% filter(is.na(spp)) %>% select(NOM_ESP, spp) %>% distinct
 # missingcodes <- anti_join(quebec.veg, taxon.info, by = c("spp" = "Code")) %>% select(NOM_ESP, spp) %>% distinct

quebec.veg <- quebec.veg %>% mutate(Cover = as.numeric(COUVERT), Species = spp, Layer = as.character(layer)) %>% select(PlotNumber, Species, Cover, Layer)
xx <- quebec.veg %>% filter(quebec.veg$Cover == 51)
#recode Cover 1 = 0.2, 2 = 1.1, 3 = 6, 4 = 16, 5 = 26, 6 = 51, 7 = 76, 9 = .1, -99 = .1
quebec.veg$Cover <- ifelse(quebec.veg$Cover == 1, 0.5, 
                           ifelse(quebec.veg$Cover == 2, 2.5, 
                                  ifelse(quebec.veg$Cover == 3, 11, 
                                         ifelse(quebec.veg$Cover == 4, 18, 
                                                ifelse(quebec.veg$Cover == 5, 38, 
                                                       ifelse(quebec.veg$Cover == 6, 63, 
                                                              ifelse(quebec.veg$Cover == 7, 88, 
                                                                     ifelse(quebec.veg$Cover == 9, 0.1, 
                                                                            ifelse(quebec.veg$Cover == -99, 0.1, quebec.veg$Cover)))))))))
duplicates <- quebec.veg[duplicated(quebec.veg) | duplicated(quebec.veg, fromLast = TRUE)]
quebec.veg2 <- quebec.veg %>% distinct(PlotNumber, Species, Layer, .keep_all = TRUE)
#quebec.veg.wide[quebec.veg.wide == 0] <- NA
## wide format
quebec.veg.wide <- quebec.veg2 %>% pivot_wider(id_cols = c("PlotNumber", "Species"), names_from = "Layer", values_from = "Cover")
# convert 0 to NA


#quebec.veg.wide <- quebec.veg %>% pivot_wider(id_cols = c("PlotNumber", "Species"), names_from = "Layer", values_from = "Cover") %>%  mutate(across(where(is.numeric), round, 1)) %>% data.frame

quebecmaster <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=E:/Arctic/Vpro64_Data_Arctic/Quebec2025.accdb;")
dbWriteTable(quebecmaster, "Quebec_2025_vegdata_long", quebec.veg2, overwrite=TRUE, batch_rows = 1)
dbDisconnect(quebecmaster)

quebec.loc <- quebec.raw %>% select(PlotNumber, Longitude, Latitude, Elev) %>% distinct 
fwrite(quebec.loc, "D:/GitHub/CanadianArctic_BGCMap/TrainingPts/Quebec_Plots_2025.csv")
pnts_sf <- st_as_sf(quebec.loc, coords=c("Longitude", "Latitude"), crs = 4326) %>% st_transform(pcrs = 3005)
mapview::mapView(pnts_sf,   cex = 6, legend = F, 
                 label = pnts_sf$PlotNumber, 
                 map.types = "OpenTopoMap")
#quebec.arctic <- fread("D:/GitHub/CanadianArctic_BGCMap/TrainingPts/Quebec_ArcticPlotID_2025_2.csv") ## generated from QGIS overlay
# quebec.arctic.raw <- quebec.raw %>% filter(NUM_PARC %in% quebec.arctic$NUM_PARC)
# quebec.raw <- left_join(plot.id,quebec.raw)
# 
# #plot.id2 <- plot.id %>% mutate(PlotNumber = str_extract(NUM_PARC, "[^_]+")) %>% mutate(nchar=nchar(PlotNumber))
# #plot.id <- fread("E:/Arctic/Quebec_data/PlotID_fixed.csv")
# 
# quebecmaster <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=E:/Arctic/Quebec_data/Quebec_Ecodata_2025.accdb;")
# dbWriteTable(quebecmaster, "Quebec_2025_ArcticPlots", quebec.raw, overwrite=TRUE, batch_rows = 1)
# dbDisconnect(quebecmaster)
```


```{r clean codes}
plot.veg$Species <- plot.veg$Species %>%  as.character() %>% toupper() %>% str_trim("right")
#plot.veg <- plot.veg %>% na_if(0) ## update zero entries to na
vegDat_nomatch1 <- anti_join(plot.veg, taxon.info) 
Error1 <- nrow(vegDat_nomatch1)

###update codes based on global code revision table (for geographically defined taxon updates)
# taxon_update <- taxon_geog_update[!taxon_geog_update$GlobalUpdate == "",] %>% distinct()
# plot.veg1 <- left_join (plot.veg,taxon_update)
# plot.veg1$Species <- ifelse(is.na(plot.veg1$GlobalUpdate), plot.veg1$Species, plot.veg1$GlobalUpdate)
# plot.veg <- plot.veg1 %>% dplyr::select(-GlobalUpdate)
# 
# vegDat_nomatch2 <- anti_join(plot.veg, taxon.info) 
# Error2 <- nrow(vegDat_nomatch2)

### update bad codes based on lookup field
# taxon_update2 <- taxon_badcode_update[!taxon_badcode_update$Code == "",] %>% distinct() %>% dplyr::rename("Species" = OldCode)
# plot.veg1 <- left_join (plot.veg,taxon_update2, by = "Species")
# plot.veg1$Species <- ifelse(is.na(plot.veg1$Code), plot.veg1$Species, plot.veg1$Code)
# plot.veg <- plot.veg1 %>% dplyr::select(-Code)
# 
# vegDat_nomatch3 <- anti_join(plot.veg, taxon.info) 
# Error3 <- nrow(vegDat_nomatch3)

###update synonym codes to new codes
#taxon.all <- taxon.all %>% mutate(OldCode = OldCodeVPro)
plot.veg <- update_taxa(plot.veg, taxon.all)

#noMatch <- taxon.xwalk[taxon.xwalk$OldCodeVPro != taxon.all$Code, 5:6] %>% dplyr::rename("Species" = OldCode) ##creates code update crosswalk
not_in_national <- anti_join (plot.veg, taxon.all, by = c("Species" = "Code")) ##join tables on Species field
temp <- not_in_national %>% group_by(Species) %>% summarize(n=n())
in.allcode <- inner_join(temp, taxon.all1, by = c("Species" = "Code"))
in.pasl <- inner_join(temp, pasl, by = c("Species" = "LETTERCODE")) %>% filter(SYNONYM == "FALSE")
#in.arctic <- inner_join(temp, arctic, by = c("Species" = "LETTERCODE"))

fwrite(temp, "badcodes.csv")
fwrite(in.pasl, "update_to_national_list.csv")

temp$Species <- ifelse(is.na(temp$Code), temp$Species, temp$Code) ## update the Species field with new Code
plot.veg <- temp %>% dplyr::select(-Code)##remove Code field from original data

vegDat_nomatch4 <- anti_join(plot.veg, taxon.info) 
Error4 <- nrow(vegDat_nomatch4)

vegDat_nomatch <- anti_join(plot.veg, taxon.info) ## produce table where data does not have a matching BC code
badcodes <- vegDat_nomatch %>% select(Species) %>% distinct

# vegfields <- plot.veg %>% select(PlotNumber, Species, Layer, Collected) %>% distinct(PlotNumber, Species, Collected)
# plot.veg4 <- left_join(plot.veg3, vegfields)
# xx <- plot.veg4 %>% 
#   group_by(PlotNumber, Species) %>% 
#   filter(n()>1)


vegDat_badcode <- anti_join(vegDat_nomatch, taxon.excluded) ## produce table with unrecognized codes
fwrite(vegDat_badcode, "./out_data/ArcticMaster2024_BadSppCodes.csv", row.names = FALSE, col.names = TRUE)
### identify good species but excluded from BC
vegDat_excluded <- right_join(vegDat_nomatch, taxon.excluded)
spp_excluded <- unique(vegDat_excluded$Species)## produce table where the code is good but not in BC (supposedly)
fwrite(vegDat_excluded, "./out_data/ArcticMaster_Excluded_SppCodes.csv", row.names = FALSE, col.names = TRUE)

vegDat_nomatch5 <- anti_join(vegDat_nomatch, vegDat_excluded)
Error5 <- nrow(vegDat_nomatch5)

#fwrite(vegDat_nomatch5, "./out_data/Sadler_AllUnmatchedCodes.csv", row.names = FALSE, col.names = TRUE)
bad.codes <- vegDat_nomatch5 %>% dplyr::select(Species) %>% dplyr::count(Species)
fwrite(bad.codes, "./out_data/Bad_Codes_tobe_Updated_arctic.csv")
```


CAVA to PASL
```{r}
###add species names from National list
plot.veg2 <- plot.veg %>% select(PlotNumber, Species) %>% left_join(taxon.info) %>% select(PlotNumber, ScientificName)
PASL_missed <- anti_join(plot.veg2, taxon.info) %>% count(ScientificName)
PASL_has <- inner_join(plot.veg2, taxon.info) %>% count(ScientificName)

plot.veg3 <- plot.veg %>% select(PlotNumber, Species) %>% left_join(taxon.info) %>% filter(Lifeform < 9 | Lifeform == 12) %>% select(PlotNumber, ScientificName)
vascan_missed <- anti_join(plot.veg3, vascan.info) %>% count(ScientificName)

```


```{r duplicates}
### identify duplicate entries

veg_dups <- as.data.table(plot.veg) %>% dplyr::select(PlotNumber, Species) %>% group_by(PlotNumber, Species) %>% filter(n()>1) ## search for duplicate spp entries in plots
fwrite(veg_dups, "./out_data/Duplicate_SppCodes.csv", row.names = FALSE, col.names = TRUE)

### identify species covers entered into the wrong strata
temp2 <- taxon.info %>% dplyr::select(Code, Lifeform) %>% dplyr::rename("Species" = Code)
veg_layer <- left_join(plot.veg, temp2)



trees <- c(0,1,2); shrubs <- c(0,1,2,3,4); herbs <- c(0, 5, 6, 7, 8, 12); mosses <- c(0, 9, 10, 11,15);seedling <- c(1,2); seedling2 <- c(3,4)
vegA <- veg_layer %>% dplyr::select(PlotNumber, Species, TotalA, Lifeform) %>% filter(!is.na(TotalA)) %>% filter(!Lifeform %in% trees) %>% drop_na()
vegB <- veg_layer %>% dplyr::select(PlotNumber, Species, TotalB, Lifeform) %>% filter(!is.na(TotalB)) %>% filter(!Lifeform %in% shrubs) %>% drop_na()
vegC <- veg_layer %>% dplyr::select(PlotNumber, Species, Cover6, Lifeform) %>% filter(!is.na(Cover6)) %>% filter(!Lifeform %in% herbs) %>% drop_na()

### add new entry for trees as mosses with cover in moss layer and remove entry in tree layer
veg_seedling <-  veg_layer %>% filter(!is.na(Cover7)) %>% filter(Lifeform %in% seedling) %>% drop_na(Lifeform) %>% as.data.table()
veg_seedling <-as.data.table(veg_seedling)


## if there is cover 5 values then simply delete the cover 7 value
#veg_seedling <- veg_seedling[veg_seedling[,c(4:25)] > 0, c(4:25) := NA]
### move shrub seedlings to B layer and remove
veg_seedling2 <-  veg_layer %>% dplyr::select(PlotNumber, Species, Cover7, Lifeform) %>% filter(!is.na(Cover7)) %>% filter(Lifeform %in% seedling2) %>% drop_na(Lifeform)

vegD <- veg_layer %>% dplyr::select(PlotNumber, Species, Cover7, Lifeform) %>% filter(!is.na(Cover7)) %>% filter(!Lifeform %in% mosses) %>% drop_na()
```

##Data fixes
Add total A and total B where not in data
Add A2 and B2 where totals only
```{r update species strata totals}
pc <- plot.veg
# update total columns if no value
pc$TotalA[!(is.na (pc$TotalA))] <- NA
pc$TotalA[pc$TotalA == 0] <- NA# sets any total that was entered as zero to NA
pc$TotalA <- ifelse(is.na(pc$TotalA), rowSums (pc[, c(4,6,8)], na.rm = TRUE), NA) # where covers 1,2,3 sum >1 put in total where no totals exist
pc$TotalA[pc$TotalA == 0] <- NA
pc$TotalA[pc$TotalA >99] <- 99

pc$TotalB[!(is.na (pc$TotalB))] <- NA # sets any total that was entered as zero to NA
pc$TotalB[pc$TotalB == 0] <- NA
pc$TotalB <- ifelse(is.na(pc$TotalB), rowSums (pc[,c(12, 14, 16, 18, 20)], na.rm=TRUE), NA)
pc$TotalB[pc$TotalB == 0] <- NA
pc$TotalB[pc$TotalB >99] <- 99
## add the total to A2 or B2 where only totals have been entered

pc$Cover2 <- ifelse(pc$TotalA > 0 & rowSums (pc[,c(4,6,8)], na.rm=TRUE) == 0, pc$TotalA, pc$Cover2)
pc$Cover5 <- ifelse(!(is.na(pc$TotalB)) & rowSums (pc[, 12:21], na.rm=TRUE) == 0, pc$TotalB, pc$Cover5)
## Checks to see where problems are
# nototA <- pc %>%
#     mutate(Total = select(., c(Cover1,  Cover2, Cover3)) %>% rowSums(na.rm = TRUE)) %>% select (PlotNumber,Species,Total, TotalA) %>% filter(Total > 0, is.na(TotalA))
# # 
# onlytotA <- pc %>%
#     mutate(Total = select(., c(Cover1, Cover2, Cover3)) %>% rowSums(na.rm = TRUE)) %>% select (PlotNumber,Species,Total, TotalA) %>% filter(Total == 0, !(is.na(TotalA)))
# # pc[paste0(pc$PlotNumber,  pc$Species) %in% paste0(notot$PlotNumber, notot$Species),] <- 
#plot.veg <- pc %>% na_if(0)
veg_dups2 <- as.data.table(plot.veg) %>% dplyr::select(-Layer, -Collected, -Flag, -ID) %>% mutate(HeightB = as.numeric(HeightB)) #group_by(PlotNumber, Species) %>% filter(n()>1)
plot.veg2  <- veg_dups2 [, lapply(.SD, sum, na.rm = TRUE), by = c("PlotNumber", "Species")]

plot.veg2[plot.veg2 == 0] <- NA
vegfields <- plot.veg %>% select(PlotNumber, Species, Layer, Collected) %>% distinct(PlotNumber, Species, Collected)
plot.veg4 <- left_join(plot.veg2, vegfields)
# xx <- plot.veg4 %>% 
#   group_by(PlotNumber, Species) %>% 
#   filter(n()>1)

### odbc export _VEG ACCESS table for updating in VPRO is too slow and fails on complete table so not used
#plot.veg4 <- plot.veg4 %>% select(-n) 
xx <- plot.veg4 %>% filter(Species == "PSEUMEN2")
plot.veg5 <- plot.veg4 %>% mutate(PlotNumber = as.character(PlotNumber)) %>% dplyr::select(-n) #arrange(desc(PlotNumber)) %>% 
# %>% #  mutate(across(where(is.numeric), round, 2)) %>% as.data.table
plot.veg5 <- setDT(plot.veg5)[order(PlotNumber, decreasing = TRUE)]
fwrite(plot.veg5, "./clean_data/BECMaster_Veg_updates.csv")


#  fieldnames = colnames(plot.veg5) 
#  vegupdate <- dbConnect(odbc::odbc(), .connection_string = "Driver={Microsoft Access Driver (*.mdb, *.accdb)}; DBQ=D:/BECMaster64/vegupdate.accdb;")
# dbCreateTable(vegupdate, "BECMaster64_updated_Veg", fields = fieldnames)
# dbDisconnect(vegupdate)


```
## export a table of Plot|Species|Type|Cover for classification work
```{r R-export}
##sum across Total A, TotalB, 6, 7, 
veg_anal <- plot.veg4 %>% dplyr::select (PlotNumber, Species, TotalA, TotalB, Cover6, Cover7, Cover8, Cover9, Cover10)  %>%  #%>% replace(is.na(.), 0)
   mutate(Cover = rowSums(.[,3:9], na.rm = T))
veg_anal <- veg_anal %>% left_join(taxon.info, by = "Species") %>% dplyr::select(PlotNumber, Species, Cover, Lifeform) #%>% na_if(0)
fwrite(veg_anal, "./clean_data/BECMaster_VegR_clean.csv", row.names = FALSE, col.names = TRUE)
## left join to species list

```

```{r update layer totals}
strat.tot <- setDT(plot.veg4)[, list(TotA = sum(TotalA, na.rm = TRUE),TotB = sum(TotalB, na.rm = TRUE),TotC = sum(Cover6, na.rm = TRUE), TotD = sum(Cover7, na.rm = TRUE)), by=PlotNumber]
strat.tot[,2:5][strat.tot[,2:5] > 100] <- 100
fwrite(strat.tot, "./clean_data/strata_totals_calculated.csv")
veg.layer <- plot.env %>% select(PlotNumber, StrataCoverTree, StrataCoverShrub, StrataCoverHerb, StrataCoverMoss)
veg.layer2 <- veg.layer %>% filter_at(vars(StrataCoverTree, StrataCoverShrub, StrataCoverHerb, StrataCoverMoss),any_vars(!is.na(.)))
veg.layer2[,2:5][veg.layer2[,2:5] > 100] <- 100
veg.layer2[is.na(veg.layer2)]  <- 100

veg.strat.compare <- left_join(veg.layer2, strat.tot) %>% mutate(StrataCoverTree = ifelse(StrataCoverTree == 0 & TotA >0, TotA, StrataCoverTree),
                                                                 StrataCoverShrub = ifelse(StrataCoverShrub == 0 & TotB >0, TotB, StrataCoverShrub),
                                                                 StrataCoverHerb = ifelse(StrataCoverHerb == 0 & TotC >0, TotC, StrataCoverHerb),
                                                                 StrataCoverMoss = ifelse(StrataCoverMoss == 0 & TotD >0, TotD, StrataCoverMoss))
veg.add.strat <- anti_join(veg.layer, veg.layer2, by = "PlotNumber")

veg.add.strat<- left_join(veg.add.strat, strat.tot) %>% mutate(StrataCoverTree =  TotA,
                                                                 StrataCoverShrub =  TotB,
                                                                 StrataCoverHerb =  TotC,
                                                                 StrataCoverMoss = TotD)
veg.layer.new <- rbind(veg.strat.compare,veg.add.strat)

fwrite(veg.layer.new, "./clean_data/strata_totals_calculated.csv")                   
###total veg cover for admin table
veg.tot <- veg.layer %>% mutate(StrataCoverTotal_calc = rowSums(.[,2:5], na.rm = T)) %>% mutate(StrataCoverTotal_calc = ifelse(StrataCoverTotal_calc>100, 100, StrataCoverTotal_calc))
tot.tot <- plot.info %>% select(Plot, StrataCoverTotal)
setDT(tot.tot)[veg.tot, StrataCoverTotal := StrataCoverTotal_calc, on = c(Plot = "PlotNumber")]

fwrite(tot.tot, "./clean_data/total_vegcover_calculated.csv")


```



```{r flag plots with high unknown/genus/family level covers}
### plots with >X cover of low specificity get low grade in veg quality rankings

```


```{r summary stats}

##############Some stats on imported data
####Count number of columns (species) required
### should use some stats from the Power of BEC script for lifeform counts
# 
# veg_exc <- vegDat_excluded %>% dplyr::select(Species) %>% group_by(Species) %>% tally() %>% as.data.frame()
# ggplot(veg_exc)+
#    geom_qq(aes(sample = n))
# 
# veg_bad <- vegDat_badcode  %>% dplyr::select(Species) %>% group_by(Species) %>% tally() %>% as.data.frame()
# ggplot(veg_bad, aes(x=reorder(Species, -n), y=n))+
#    geom_point()
# 
# veg_count <- plot.veg  %>% dplyr::select(Species) %>% group_by(Species) %>% tally() %>% as.data.frame()
# ggplot(veg_count)+
#    geom_qq(aes(sample = n))
# 
# veg_norare <- veg_count[!veg_count$n <=5,]
# ggplot(veg_norare)+
#    geom_qq(aes(sample = n))
# ###Plot number of species in density graph
# ggplot(data=veg_norare,aes(x=reorder(Species, -n), y=n))+
#  geom_point()


# ggplot(plot.veg,aes(Species))+
#   geom_density(adjust = 5)

```

SadlerThesis_CleanedData_VascularMatrix

```{r Sadler data temporary import}
# sadler1 <- fread("./raw_data/SadlerThesis_CleanedData_VascularMatrix.csv") %>% 
#   select(-Type, -Site, -Plot, -FieldNumber) %>% pivot_longer(-PlotNumber, names_to = "Species", values_to = "cover") %>% mutate(layer = "6")
# sadler2 <- fread("./raw_data/SadlerThesis_CleanedData_CryptoMatrix.csv") %>% 
#   select(-MPlot, -Site, -Plot, -FieldNumber) %>% pivot_longer(-PlotNumber, names_to = "Species", values_to = "cover") %>% mutate(layer = "7")
# sadler <- rbind(sadler1,sadler2) %>% drop_na()# %>% filter(!Code == "")
# plot.veg = sadler
```
